---
title: "Project_607"
author: "Andrianina David Raharijao"
date: "2024-05-06"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(dplyr)
library(ggplot2)
library(readr)
library(ggmap)
```

##Data acquisition

**CSV data of MTA Subway Stations and their location**
```{r read_data}
mta_stations <- read_csv("https://raw.githubusercontent.com/Raharijao/Spring2024_Data_607_CUNYSPS/main/Project_final/MTA_Subway_Stations_20240506.csv")
```

**API call of ridership data**
```{r api_call_1}
url <- "https://data.ny.gov/resource/wujg-7c2s.json"
app_token <- "85966iMGG1JQ0cq5w2rJmYTxQ"
```

```{r api_call_2}
response <- GET(url, add_headers(`X-App-Token` = app_token))
```

**Pull the API data, only bull subway info for Manhattan and the Bronx, and pick the hourly data for Monday April 22nd, at 8am**
```{r api_call_3}
# Setup query parameters for the API call
query_params <- list(
  `$where` = paste("transit_mode='subway' AND",
                   "(borough='Manhattan' OR borough='Bronx') AND",
                   "transit_timestamp >= '2024-04-25T08:00:00'"),
  `$limit` = 2000,
  `$$app_token` = "85966iMGG1JQ0cq5w2rJmYTxQ"
)
response2 <- GET(url, query = query_params)

data <- content(response2, "parsed")
data_frame <- bind_rows(data)
```

##Data Cleaning

1- CSV data on Train stations

**Renamed a column**
```{r cleaning_1_a}
mta_stations <- mta_stations %>%
  rename(train_Line = `Daytime Routes`) %>%
  rename(station_complex_id = `Complex ID`)%>%
  mutate(
    station_complex_id = as.character(station_complex_id))
```

**Only kept stations that have the D train, and only kept Manhattan and Bronx stops, and delete some columns**
```{r cleaning_1_b}
mta_stations_D <- mta_stations %>%
  filter(grepl("D", train_Line)) %>%
  filter(Borough %in% c("M", "Bx")) %>%
  select(-Division, -Structure, -ADA, -`ADA Southbound`,-`ADA Northbound`,-`North Direction Label`,-`South Direction Label`, -`ADA Notes`)
```

2- Data from API call
**Removing duplicates, converting the transit time stamp column to a time value, and the ridership data to int**
```{r cleaning_2_a}
data_frame_clean <- data_frame %>%
  filter(georeference != "Point") %>%
  select(transit_timestamp, station_complex_id, payment_method, fare_class_category, ridership, transfers)%>%
  mutate(transit_timestamp = as.POSIXct(transit_timestamp, format="%Y-%m-%dT%H:%M:%S", tz="GMT"))%>%
arrange(desc(transit_timestamp))%>%
  mutate(
    ridership = as.integer(ridership),
    transfers = as.integer(transfers))
```


##Merged Data:
**Merge the two data set on the station complex ID**
```{r merge_1}
merged_data <- inner_join(data_frame_clean, mta_stations_D, by = "station_complex_id", suffix = c("", "_D")) %>%
  select(everything(), 'Stop Name', 'GTFS Latitude', 'GTFS Longitude')
```
**Remove unwanted columns**
```{r merged_2}
merged_data_clean <- select(merged_data, 
                              -transit_timestamp,
                              -Line,
                              -Borough,
                              -train_Line,
                              -station_complex_id, 
                              -`GTFS Stop ID`, 
                              -`Station ID`, 
                              -Georeference)
```
**Add ridership and transfer**
```{r merged_3}
merged_data_clean <- merged_data_clean %>%
  mutate(total_ridership = ridership + transfers)%>%
  select(-ridership, -transfers)%>%
  rename(Stop_name = `Stop Name`)%>%
  rename(GTFS_Latitude = `GTFS Latitude`)%>%
  rename(GTFS_Longitude = `GTFS Longitude`)
```

##Statistical analysis and graphics.
**Summary statistics on which fare type is the most popular one**
```{r stat_1}
payment_method_summary <- merged_data_clean %>%
  group_by(fare_class_category) %>%
  summarise(Total_Ridership = sum(total_ridership, na.rm = TRUE),
            Average_Ridership = mean(total_ridership, na.rm = TRUE),
            Count = n()) %>%
  arrange(desc(Total_Ridership))
print(payment_method_summary)

```

**Rideship by stop**
```{r visual_1}
plot_data <- merged_data_clean %>%
  group_by(Stop_name, GTFS_Latitude, GTFS_Longitude) %>%
  summarise(total_ridership = sum(total_ridership, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(GTFS_Latitude))
```

```{r visual_2}
ggplot(plot_data, aes(x = reorder(Stop_name, total_ridership), y = total_ridership)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Total Ridership by Subway Stop",
       x = "Subway Stop",
       y = "Total Ridership") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

##Using ggmap to plot out the result:

**Set up the stadia map of NYC**
```{r map_1}
#API credentials had to first be set ip
register_stadiamaps(key = "fdfed2b0-ab6e-4805-bff6-f450ea8b212e")
bbox <- c(left = -74.05, bottom = 40.68, right = -73.85, top = 40.85)
nyc_map <- get_stadiamap(bbox = bbox, zoom = 12, maptype = "stamen_terrain_lines")

```
**Plot the ridership data by station on the map**
```{r map_2}
ggmap(nyc_map) +
  geom_point(data = plot_data, aes(x = GTFS_Longitude, y = GTFS_Latitude, size = total_ridership, color = total_ridership), alpha = 0.5) +
  geom_line(data = plot_data, aes(x = GTFS_Longitude, y = GTFS_Latitude, group = 1), color = "orange", size = 1) +
  scale_size_continuous(range = c(1, 10)) + 
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Thursday April 25nd 8am", x = "Longitude", y = "Latitude") +
  theme_minimal()+
  theme(legend.position = "none") 

```

