---
title: "Assignment 1"
author: "Andrianina David Raharijao"
date: "2024-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Introduction

For this assignment, I have decided to use the "What Do Men Think It Means To Be A Man?" data set. I will be looking at the raw responses.

Each row represents one respondent and each column is the question that they answered, and the row content contains their answers to the question.

The [article](https://fivethirtyeight.com/features/what-do-men-think-it-means-to-be-a-man/) related to the data set was written by Ella Koeze and Anna Maria Barry Jester. It reports on a survey of 1,615 American men, and talks about how their perception on masculinity changed after the #MeToo movement. 

**My subset of the data frame**

I found it interesting that some respondent of the respondent deemed themselves as "Very masculine", and so I would like to know more about the subgroup of people that answered the first question this way.

I will be keeping the following column:

* q0001: How masculine do you feel ?
* q0002: How important is other's perception of your masculinity?
* q0026: Sexual orientation
* q0029: last grade of school completed
* age3:  age range

```{r Q1_a}
#Load the csv file
#Old way
#response <- read_csv("raw-responses.csv")

#New way
url <- https://raw.githubusercontent.com/Raharijao/Spring2024_Data_607_CUNYSPS/main/Week1/raw-responses.csv
response <- read_csv(url)
```
```{r Q1_b}
#Looking at the initial table
response

```
```{r remove other columns}
#Only keeping the questions that I want to observe
cl <- response[, c("q0001","q0002","q0026","q0029","age3")]

```
```{r add column}
#Adding an index to distinguish each respondents
cl$respondent_id <- seq.int(nrow(cl))
```
```{r rename columns}
#Rename columns
cl <- rename(cl,
   masculinity_lvl = q0001 ,
   masculinity_importance = q0002,
   sex_orientation = q0026,
   highest_ed = q0029,
   age_group = age3
  )
```
```{r modify columns}
#Modify column content
cl$masculinity_lvl <- gsub("masculine", "", cl$masculinity_lvl, ignore.case = TRUE)
cl$masculinity_importance <- gsub("important", "", cl$masculinity_importance, ignore.case = TRUE)
```
```{r filter}
#Only filter out my desired rows
cl_v <- cl %>%
  mutate(across(where(is.character), trimws))

cl_v <- cl_v %>%
  filter(masculinity_lvl == "Very")
```
```{r table}
cl_v
```

## Conclusion

From this selected article, I intend to explore the data to see if there is a strong correlation with the respondent putting "Very masculine" and one of the other variables that I have kept in the table. This is to see if a specific background can be attributed to a person's image of their masculinity.

