---
title: "Project_1"
author: "Andrianina David Raharijao"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(stringr)
library(tidyr)
```

## Planning

First I want to determine the columns that I will need in the final csv file:

* Player name
* Player state
* Total number of points
* Player Pre-rating
*Average Pre Chess
  *Opponents pre tournament ratings (sum)
  *Total number of opponents

# Loading the data
```{r load_data}
data <- readLines("Chess_data.txt")

```

#Extracting the player number
```{r Player_numb}
pnu_pattern <- "\\d+ \\| "
pnu <- str_extract_all(data, pnu_pattern)
pnu <- unlist(pnu)
pnu <- gsub(' \\| ', "", pnu)
print(pnu)
```

#Extracting the player's name
```{r Player_name}
pn_pattern <- "\\| [A-Z]([^|]+) \\|"
pn <- str_extract_all(data, pn_pattern)
pn <- unlist(pn)
#remove extra characters
pn <- gsub('\\|', "", pn)
#remove header rows
pn <- pn[-c(1,2)]
#remove extra spaces
pn <- trimws(pn)
print(pn)
```

#Extracting the player's state
```{r Player_state}
ps_pattern <- "[A-Z][A-Z] \\|"
ps <- str_extract_all(data, ps_pattern)
ps <- unlist(ps)
ps <- gsub(' \\|', "", ps)
print(ps)
```

#Extracting the player's total point
```{r Player_total_point}
pts_pattern <- "\\d\\.\\d+  \\|"
ptS <- str_extract_all(data, pts_pattern)
ptS <- unlist(ptS)
ptS <- gsub('  \\|', "", ptS)
print(ptS)
```

#Extracting the player's pre chess rating
```{r Player_pre_chess_rating}
pcrt_pattern <- "R: ([^-]+)-"
pcr_pattern <- "R:\\s*(\\d{1,4})"
pcr <- str_extract_all(data, pcr_pattern)
pcr <- unlist(pcr)
pcr <- gsub('R: ', "", pcr)
#pcr <- gsub('-', "", pcr)
pcr <- trimws(pcr)
print(pcr)
```

#Building the data frame (to be used in next step)
```{r dataframe1}

df <- data.frame(pn,ps,ptS,pcr)
df$pcr <- as.integer(df$pcr)

```

#Extracting a list of all the oppopnents for each row
```{r Player_opponents}
pop_pattern <- "[A-Z] \\s*\\d+"
pop <- str_extract_all(data, pop_pattern)
#Filters out blank lists with no opponent info
pop <- pop[lengths(pop) > 0]
#Clean out non player ID characters
pop <- lapply(pop, function(x) gsub("[A-Z]  ", "", x))
# #Make it into a list of number lists
pop <- lapply(pop, unlist)
pop <- lapply(pop, function(sublist) {
   as.numeric(sublist)
 })
print(pop)
```

#Extracting all of the opponent's average score from the data frame and then averaging it
```{r average_opponents_score}

#For each sublist in the main list
popt <- lapply(pop, function(numb_list) {
  #For each items in the sublist
  t_scores <- lapply(numb_list, function(number) {
    if (number %in% rownames(df)) {
      return(df[number, "pcr"])
    } else {
      return(NA)
    }
  })
  av <- round(mean(unlist(t_scores)))
  return(av)
})
popt <- unlist(popt)
print(popt)


```

#Building final data frame
```{r dataframe2}

df$popt <- as.integer(popt)

#Rename columns
df <- rename(df,
   Player_name = pn ,
   Player_state = ps,
   Player_total_points = ptS,
   Player_pre_rating = pcr,
   Player_op_avg_prerat = popt
  )

```


These are just chunks of codes I used to experiment
```{r test}
data_test <- c("aaa bfg gl","klmno jjj", "kop lll ooo")
data_test2 <- c("aaa", "abc", "bbbbb", "cccd", "eeef", "ggghhh")
pattern <- "(.)\\1\\1"
matching_elements <- str_extract_all(data_test, pattern)
matching_elements <- unlist(matching_elements)
print(matching_elements)

```

```{r test2}
test <- "|6.0  |W  39|W  21|W  18|W  14|W   7|D  12|D   4|"
pop_pattern <- "[A-Z]\\s*\\d+"
testr <- str_extract_all(test, pop_pattern)
print(testr)


```

```{r test3}
data_test <- c("R: 123lkj")
pattern <- "R: (\\d{1,4})" 
matching_elements <- str_extract_all(data_test, pattern)
matching_elements <- unlist(matching_elements)
print(matching_elements)

```