---
title: "Assignment 3"
author: "Andrianina David Raharijao"
date: "2024-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
library(tidyverse)
#Setting up data set
majors <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
```

## Question 1

```{r q1}
q1_ans <- majors %>%
  filter(grepl("DATA", Major) | grepl("STATISTICS", Major))
q1_ans
```

## Question 2

```{r q2}
#Loads fruit list
fruit_list <- list("bell pepper",  "bilberry",     "blackberry",   "blood orange",
 "blueberry",    "cantaloupe",   "chili pepper", "cloudberry",  
 "elderberry",   "lime",         "lychee",       "mulberry",    
 "olive",        "salal berry")

# Turns the list info a vector
q2_ans <- unlist(fruit_list)
q2_ans
```
## Question 3

* (.)\1\1 = matches any sequence of a single character repeating three times. (ex: "aaa")
* "(.)(.)\\2\\1" = matches any sequences of characters that look like this (abba, cddc, effe) and is in between ""
* (..)\1 = matches any sequences of characters that look like this (abab, cdcd, efef)
* "(.).\\1.\\1" = matches any sequences of characters that look like this (adaxa, ckchc, nlnjn, lmlnl) and is between ""
* "(.)(.)(.).*\\3\\2\\1" = matches any sequences of characters where it begins with a set of three characters, followed by some character (or none)and then ending with the first set of three characters, but in reverse order. All of this between ""

```{r q3}
#For testing
text <- 'abccbalkjsdoihfoie'
pattern1 <- '(.)(.)(.).*\\3\\2\\1'

q3_ans <- regmatches(text, gregexpr(pattern1, text))
q3_ans
```

## Question 4
```{r q4}
#For testing
text <- 'lkjhklk lilieii hkihiii iii'

regex_1 <- "\\b(\\w)\\w*\\1\\b"
#//b denotes the limits of the word
#(\\w) denotes the first character of the word
#\\w* represents all following character
#\\1 means that the last character should be the same as the first character
regex_2 <- "\\b\\w*(\\w\\w)\\w*\\1\\w*\\b"
#//b denotes the limits of the word
#\\w* denotes any character (or none) before the two sets of character
#(\\w\\w) represents the two characters
regex_3 <- "\\b(\\w).*\\1.*\\1.*\\b"

q4_ans <- regmatches(text, gregexpr(regex_3, text))
q4_ans
```
